Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-01 prop_letter.tex
prop_letter.tex \section{Kochen-Specker} Updating...

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality~\cite{BarrettKent2004,Appleby_2005}. The outcome
of a theory that is contextual depends upon
whether  compatible sets of observables are measured together or
separately; a non-contextual theory gives the same results in either
case.  A classic example of the varying opinions on the impact of
imprecision is the disagreement between
Meyer~\cite{PhysRevLett.83.3751} and 
Mermin~\cite{Mermin1999}  on whether finite-precision
measurements invalidate the spirit of the Kochen-Specker
theorem~\cite{kochenspecker1967,peres1995quantum}.  The Kochen-Specker  
theorem is in essence a
mathematical statement about contextuality, asserting that
in a Hilbert space of dimension $d \ge 3$, it is
impossible to associate definite measured values, $\mu(P_i)= 0$ or
$1$, with every projection operator~$P_i$, in such a way that, if a
set of commuting $P_i$ satisfies $\sum_i P_i = 1$, then
$\sum_i \mu(P_i) = 1$.   Meyer showed that one {\it can\/}
assign definite values $0$ or $1$ to the measurement outcome
$\mu(P_i)$ when the Hilbert space is defined over the field of
rational numbers, therefore nullifying the
theorem~\citep{PhysRevLett.83.3751}.  However, Mermin then argued
that measurement outcomes must depend smoothly on slight changes in
the experimental configuration, leading him to assert that the
impact of Meyer's negative result is
``unsupportable''~\citep{Mermin1999}.  The debate on
whether contextuality is consistent with experimental
outcomes continues to be an active topic of research~\citep{Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

@@

Our generalization of quantum probability measures to QIVPMs allows us
to strengthen the scope of one of the fundamental theorems of quantum
information: the Kochen-Specker
theorem~\cite{PeresRon1988,peres1995quantum}. Our finite-precision
extension of the theorem will suggest a resolution to the debate
initiated by Meyer and Mermin on the relevance of the Kochen-Specker
to experimental, and hence finite-precision, quantum
measurements~\citep{PhysRevLett.83.3751,Mermin1999,Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.
Specifically, the original Kochen-Specker theorem is formulated under
the assumption that all observables of a quantum mechanical system
have \emph{definite values at all times}~\cite{Held2016}. Our
interval-valued probability framework will allow us to state, and
prove, a stronger version of the theorem that holds even if the
observables have values that are only definite up to some specific
parameter that quantifies experimental uncertainty. Put differently,
our stronger statement and theorem quantifies Mermin's
intuition~\citep{Mermin1999}:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case.
\end{quote}

@@

When $\delta=0$ , every event must be assigned a probability in either
the interval $[0,0]$ or $[1,1]$, i.e., every event is completely
determined with certainty. As $\delta$ gets bigger, the QIVPM allows
more and more indeterminate behavior. Our technical result will be
that for ``small enough'' $\delta$, a variant of the Kochen-Specker
theorem is still valid. To state and prove this result we will first
reformulate the original Kochen-Specker as a statement regarding the
non-existence of $0$-deterministic QIVPM. We will then strengthen this
result to show the non-existence of any $\delta$-deterministic QIVPM
when $\delta < 1/3$ and the Hilbert space dimension is greater than
2.

@@

A 0-deterministic QIVPM\nb$\bar{\mu}$ has an important property
regarding to the expectation values of a family of commuting observables
relative to $\bar{\mu}$. A family of $M$ commuting observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{M}$
can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{k}\right\} _{k=0}^{d-1}$.
Since $\bar{\mu}$ is 0-deterministic, a unique $\ket{k_{0}}$ exists
such that $\bar{\mu}\left(\proj{k}\right)=\necess$ if and only if
$k=k_{0}$\nb\citep{HOSTunpublished}. Also, a unique quantum probability
measure\nb$\mu$ exists such that $\bar{\mu}\left(P\right)=\left\{ \mu\left(P\right)\right\} $
and $\mu\left(P\right)\in\left\{ 0,1\right\} $ for all $P\in\events$.
Especially, $\mu\left(\proj{k}\right)=1$ if and only if $k=k_{0}$.
If we denote the product observables\nb$\prod_{j=1}^{M}\mathbf{O}_{j}$
as $\overline{\mathbf{O}}$, the expectation values of the observables\nb$\expval{\mathbf{O}_{j}}_{\bar{\mu}}$
and $\expval{\overline{\mathbf{O}}}_{\bar{\mu}}$ are equal to the
singleton sets whose only elements are $\expval{\mathbf{O}_{j}}_{\mu}$
and $\expval{\overline{\mathbf{O}}}_{\mu}$, respectively. Both $\prod_{j=1}^{M}\expval{\mathbf{O}_{j}}_{\mu}$
and $\expval{\overline{\mathbf{O}}}_{\mu}$ are equal to the product
of the eigenvalues corresponding to the pure state\nb$\ket{k_{0}}$
for each observable\nb$\mathbf{O}_{j}$. By adopting an intuitive
product rule\nb$\left\{ x_{0}\right\} \cdot\left\{ x_{1}\right\} =\left\{ x_{0}x_{1}\right\} $,
we have the following property: 
\begin{equation}
\prod_{j=1}^{M}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{M}\mathbf{O}_{j}}_{\bar{\mu}}\,.\label{eq:Kochen-Specker-Rules}
\end{equation}

>>>>

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality~\cite{BarrettKent2004,Appleby_2005}. The outcome
of a theory that is contextual depends upon
whether  compatible sets of observables are measured together or
separately; a non-contextual theory gives the same results in either
case.  A classic example of the varying opinions on the impact of
imprecision is the disagreement between
Meyer~\cite{PhysRevLett.83.3751} and 
Mermin~\cite{Mermin1999}  on whether finite-precision
measurements invalidate the spirit of the Kochen-Specker
theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007,Held2016}.  The Kochen-Specker  
theorem is in essence a
mathematical statement about contextuality, asserting that
in a Hilbert space of dimension $d \ge 3$, it is
impossible to associate definite measured values, $\mu(P_i)= 0$ or
$1$, with every projection operator~$P_i$, in such a way that, if a
set of commuting $P_i$ satisfies $\sum_i P_i = 1$, then
$\sum_i \mu(P_i) = 1$.   Meyer showed that one {\it can\/}
assign definite values $0$ or $1$ to the measurement outcome
$\mu(P_i)$ when the Hilbert space is defined over the field of
rational numbers, therefore nullifying the
theorem~\citep{PhysRevLett.83.3751}.  However, Mermin then argued
that measurement outcomes must depend smoothly on slight changes in
the experimental configuration, leading him to assert that the
impact of Meyer's negative result is
``unsupportable''~\citep{Mermin1999}.  The debate on
whether contextuality is consistent with experimental
outcomes continues to be an active topic of research~\citep{Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

@@

Our generalization of quantum probability measures to QIVPMs allows us
to strengthen the scope of one of the fundamental theorems of quantum
information: the Kochen-Specker
theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007,Held2016}. Our finite-precision
extension of the theorem will suggest a resolution to the debate
initiated by Meyer and Mermin on the relevance of the Kochen-Specker
to experimental, and hence finite-precision, quantum
measurements~\citep{PhysRevLett.83.3751,Mermin1999,Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.
Specifically, the original Kochen-Specker theorem is formulated under
the assumption that all observables of a quantum mechanical system
have \emph{definite values at all times}~\cite{Held2016}. Our
interval-valued probability framework will allow us to state, and
prove, a stronger version of the theorem that holds even if the
observables have values that are only definite up to some specific
parameter that quantifies experimental uncertainty. Put differently,
our stronger statement and theorem quantifies Mermin's
intuition~\citep{Mermin1999}:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case.
\end{quote}

@@

When $\delta=0$ , every event must be assigned a probability in either
the interval $[0,0]$ or $[1,1]$, i.e., every event is completely
determined with certainty. As $\delta$ gets bigger, the QIVPM allows
more and more indeterminate behavior. For example, when $\delta=\frac{1}{3}$,
two possible collection of intervals\nb$\mathscr{I}$ are $\left\{ \imposs,\left[0,\frac{1}{3}\right],\left[\frac{2}{3},1\right],\necess\right\} $
and $\left\{ \imposs,\left[\frac{1}{6},\frac{1}{6}\right],\left[\frac{1}{3},\frac{1}{3}\right],\left[\frac{2}{3},\frac{2}{3}\right],\left[\frac{5}{6},\frac{5}{6}\right],\necess\right\} $.
Our technical result will be
that for ``small enough'' $\delta$, a variant of the Kochen-Specker
theorem is still valid. To state and prove this result we will first
reformulate the original Kochen-Specker as a statement regarding the
non-existence of $0$-deterministic QIVPM. We will then strengthen this
result to show the non-existence of any $\delta$-deterministic QIVPM
when $\delta < 1/3$ and the Hilbert space dimension is greater than
2.

@@

A 0-deterministic QIVPM\nb$\bar{\mu}$ has an important property
regarding to the expectation values of a family of $M$ commuting
observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{M}$ relative
to $\bar{\mu}$. Since $\left\{ \mathbf{O}_{j}\right\} _{j=1}^{M}$
commutes, they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{k}\right\} _{k=0}^{d-1}$.
Since $\bar{\mu}$ is 0-deterministic, a unique $\ket{k_{0}}$ and
a unique quantum probability measure\nb$\mu$ exists such that $\bar{\mu}\left(P\right)=\left\{ \mu\left(P\right)\right\} $
and $\mu\left(P\right)\in\left\{ 0,1\right\} $ for all $P\in\events$;
$\mu\left(\proj{k}\right)=1$ if and only if $k=k_{0}$\nb\citep{HOSTunpublished}.
If we denote the product observables\nb$\prod_{j=1}^{M}\mathbf{O}_{j}$
as $\overline{\mathbf{O}}$, the expectation values of the observables\nb$\expval{\mathbf{O}_{j}}_{\bar{\mu}}$
and $\expval{\overline{\mathbf{O}}}_{\bar{\mu}}$ are equal to the
singleton sets whose only elements are $\expval{\mathbf{O}_{j}}_{\mu}$
and $\expval{\overline{\mathbf{O}}}_{\mu}$, respectively. Both $\prod_{j=1}^{M}\expval{\mathbf{O}_{j}}_{\mu}$
and $\expval{\overline{\mathbf{O}}}_{\mu}$ are equal to the product
of the eigenvalues corresponding to the pure state\nb$\ket{k_{0}}$
for each observable\nb$\mathbf{O}_{j}$. By adopting an intuitive
product rule\nb$\left\{ x_{0}\right\} \cdot\left\{ x_{1}\right\} =\left\{ x_{0}x_{1}\right\} $,
we have the following property: 
\begin{equation}
\prod_{j=1}^{M}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{M}\mathbf{O}_{j}}_{\bar{\mu}}\,.\label{eq:Kochen-Specker-Rules}
\end{equation}
